Kepler's Demographic Cliff: Ages of Young Stars and Planets in the Kepler Field

### 2024-04-30 paper/ms.tex
update ms

\begin{figure*}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.48\textwidth]{prot_teff_Santos19_Santos21_dquality.pdf}
		}
		%	
		%		\vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.48\textwidth]{koi_mean_prot_teff_koi_X_S19S21dquality_keepgrazing.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
		{\bf Rotation periods for Kepler stars (left) and planets
			(right)}.  The periods are from \citet{Santos_2019} and
		\citet{Santos_2021} (\citetalias{Santos_2019} and
		\citetalias{Santos_2021}).  The gray lines are ``mean fits'' from
		Figure~\ref{fig:models} to the slow rotation sequences of M67,
		Ruprecht~147, NGC~6811, Praesepe, and a trio of 120\,Myr open
		clusters.  The stellar sample shows only apparently single stars
		near the main sequence with $\log g$$>$4.2, RUWE$<$1.4,
		3800$<$$T_{\rm eff}$/K$<$6200, that are not flagged as eclipsing
		binaries.  The planet sample has the same stellar cuts, and shows
		confirmed and candidate planets from the cumulative KOI table,
		requiring them to have a MES$>$10.
	}
	\label{fig:prot_vs_teff}
\end{figure*}

@@

\startlongtable
\begin{deluxetable*}{llllllllrrrrrrl}
	%\tabletypesize{\footnotesize}
	\tabletypesize{\scriptsize}
  \tablecaption{Ages of Kepler planets and planet candidates.  This
  version of the table is truncated to include the youngest fifty
  confirmed Kepler planets, and the twenty youngest candidates,
  sorted by the minimum of either the rotation or lithium-based
  age.  The
  full machine-readable table contains ages and age limits for
  \nnonfopkoissomeageinfo\ non-false positive KOIs with MES$>$10.  
  \label{tab:planets}}
	%\toprule
	%\midrule
	%\endhead
	\startdata
  KOI & Kepler &  $T_{\rm eff}$ & $P_{\rm rot}$ & EW$_{\rm Li}$ &
  $t_{\rm gyro}$ & $t_{\rm Li} $ & Consistent? & $R_{\rm p}$ & $P_{\rm orb}$ &
  $Q_{\rm planet}$ & $Q_{\rm gyro}$ &  Spec? & Comment \\
  -- &   -- & K & days & m\AA & Myr & Myr &  str &   Earths &    days &       int  & int & bool & -- \\
	\hline
	\input{table_subgyr_confirmed} %\\
%	\rule
	\input{table_subgyr_candidate}
	\enddata
	\tablecomments{``Confirmed'' planets appear before``candidate'' planets.
		The bit quality flags for the rotation-based ages,
		$Q_{\rm gyro}$, are described in Section~\ref{subsec:flags}.
		Concisely summarized, they are:
		Bit 0: $T_{\rm eff}/{\rm K} \in [ 3800-6200]$ ?
		Bit 1: $\log g$$<$4.2?
		Bit 2: $M_{\rm G}$<3.9?
		Bit 3: In KEBC?
		Bit 4: Large $d_{\rm xm,Kep-Gaia}$?
		Bit 5: Confused Kep-Gaia crossmatch?
		Bit 6: Gaia DR3 non-single star?
		Bit 7: RUWE$>$1.4?
		Bit 8: Crowded?
		Bit 9: Far from main-sequence?
		Bit 10: \citetalias{Santos_2021} CP/CB?
		As an example, Kepler-1627 has $Q_{\rm gyro}$ flagged with bit 10 and bit 7.
		The analogous planet quality bit-flag, $Q_{\rm planet}$, has the
		following meaning.
		Bit 0: Candidate not reliable?
		Bit 1: MES$<$10?
		Bit 2: Grazing?
		For a star to have a high likelihood of being ``reliable for gyrochronology'' we suggest $Q_{\rm star}$ bits
		zero to nine to not be raised, and for a planet to be ``reliabe'', we suggested $Q_{\rm planet}$ to be zero.
	}
\end{deluxetable*}

>>>>

\begin{figure*}[!t]
  \begin{center}
    \subfloat{
      \includegraphics[width=0.48\textwidth]{prot_teff_Santos19_Santos21_dquality.pdf}
    }
    % 
    %   \vspace{-0.5cm}
    \subfloat{
      \includegraphics[width=0.48\textwidth]{koi_mean_prot_teff_koi_X_S19S21dquality_keepgrazing.pdf}
    }
  \end{center}
  \vspace{-0.5cm}
  \caption{
    {\bf Rotation periods for Kepler target stars (left) and known
    planet hosts (right)}.  The periods are from \citet{Santos_2019}
    and \citet{Santos_2021} (\citetalias{Santos_2019} and
    \citetalias{Santos_2021}).  The gray lines are ``mean fits'' from
    Figure~\ref{fig:models} to the slow rotation sequences of M67,
    Ruprecht~147, NGC~6811, Praesepe, and a trio of 120\,Myr open
    clusters.  The stellar sample shows only apparently single stars
    near the main sequence with $\log g$$>$4.2, RUWE$<$1.4,
    3800$<$$T_{\rm eff}$/K$<$6200, that are not flagged as eclipsing
    binaries.  The planet sample has the same stellar cuts, and shows
    confirmed and candidate planets from the cumulative KOI table,
    requiring them to have a MES$>$10.
  }
  \label{fig:prot_vs_teff}
\end{figure*}

@@

\startlongtable
\begin{deluxetable*}{llllllllrrrrrrl}
  %\tabletypesize{\footnotesize}
  \tabletypesize{\scriptsize}
  \tablecaption{Ages of Kepler planets and planet candidates.  This
  version of the table is truncated to include the youngest fifty
  confirmed Kepler planets, and the twenty youngest candidates,
  sorted by the minimum of either the rotation or lithium-based
  age.  The
  full machine-readable table contains ages and age limits for
  \nnonfopkoissomeageinfo\ non-false positive KOIs with MES$>$10.  
  \label{tab:planets}}
  %\toprule
  %\midrule
  %\endhead
  \startdata
  KOI & Kepler &  $T_{\rm eff}$ & $P_{\rm rot}$ & EW$_{\rm Li}$ &
  $t_{\rm gyro}$ & $t_{\rm Li} $ & Consistent? & $R_{\rm p}$ & $P_{\rm orb}$ &
  $Q_{\rm planet}$ & $Q_{\rm gyro}$ &  Spec? & Comment \\
  -- &   -- & K & days & m\AA & Myr & Myr &  str &   Earths &    days &       int  & int & bool & -- \\
  \hline
  \input{table_subgyr_confirmed} %\\
% \rule
  \input{table_subgyr_candidate}
  \enddata
  \tablecomments{``Confirmed'' planets appear before``candidate'' planets.
    The bit quality flags for the rotation-based ages,
    $Q_{\rm gyro}$, are described in Section~\ref{subsec:flags}.
    Concisely summarized, they are:
    Bit 0: $T_{\rm eff}/{\rm K} \in [ 3800-6200]$ ?
    Bit 1: $\log g$$<$4.2?
    Bit 2: $M_{\rm G}$<3.9?
    Bit 3: In KEBC?
    Bit 4: Large $d_{\rm xm,Kep-Gaia}$?
    Bit 5: Confused Kep-Gaia crossmatch?
    Bit 6: Gaia DR3 non-single star?
    Bit 7: RUWE$>$1.4?
    Bit 8: Crowded?
    Bit 9: Far from main-sequence?
    Bit 10: \citetalias{Santos_2021} CP/CB?
    Bit 11: $P_{\rm rot}$ not in the homogeneous
    \citetalias{Santos_2019} \citetalias{Santos_2021} sample?
    As an example, Kepler-1627 has $Q_{\rm gyro}$ flagged with bit 10 and bit 7.
    The analogous planet quality bit-flag, $Q_{\rm planet}$, has the
    following meaning.
    Bit 0: Candidate not reliable?
    Bit 1: MES$<$10?
    Bit 2: Grazing?
    For a star to have a high likelihood of being ``reliable for gyrochronology'' we suggest $Q_{\rm star}$ bits
    zero to nine to not be raised, and for a planet to be ``reliabe'', we suggested $Q_{\rm planet}$ to be zero.
  }
\end{deluxetable*}
